Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2021-01-24 manuscript/response_to_reviews.Rmd

b. I need information on how Figures 2-5, 8-11 were prepared. Carbon stocks and fluxes such as NPP, litterfall, etc. are probably plot data. There is a detailed discussion of how the data were filtered to remove, for example, disturbance effects (lines 140-149). The GPP, Reco, and soil respiration databases seem to me to be different and may not be collocated with the plot data. Was disturbance filtering applied to the GPP, Reco, and soil respiration databases, too?

**All records analyzed here have been fully incorporated into the ForC format and treated as described. We believe this should now be clearer where we describe the importation of new datasets: "Since publication of ForC v2.0, we imported three large additional databases [SRDB, GROA, FLUXNET15] into *ForC* via a combination of R scripts and manual edits. ..."**

c. If Luyssaert et al (2007) is a source of data, is there circularity in the comparison with latitudinal trends found in other datasets, for which Luyssaert et al is cited (lines 317-318)?

**We don't see this as circular; we are reaffirming earlier results with an expanded dataset. To make it clear that we are not claiming to provide an independent test, we have edited this text as follows: "For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate]. This consistency is not not surprising, particularly given commonality in the data analyzed or used for calibration."**

@@

**Thank you. We have added the following paragraph (citing the references given above) on lack of representation of fragmented and degraded forests in the Discussion section "C variable coverage and budget closure":**

**Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests--geographically or in terms of human impacts [@martin_mapping_2012]. Geographically, all variables are poorly covered in Africa and Siberia (Fig. 1), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. In terms of human impacts, research efforts tend to focus on interior forest ecosystems [@martin_mapping_2012], often in permanently protected areas [e.g., @davies_forestgeo_2021]. Studies of regrowth forests tend to focus on sites where recurring anthropogenic disturbance is not a confounding factor. Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@fao_and_unep_state_2020]. Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].Partial logging and other forms of non- stand clearing anthropogenic disturbance also strongly impact forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016] but are under-studied [@sist_tropical_2015] and excluded from this analysis. Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig. `r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019]. Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].**

@@

L322-323: As you point out in the Results section, while there are no statistically sig differences in NEP across biomes, there are large differences in the means. Is this an artifact of the data sources used (i.e. distribution and number of sites with empirical data)? Can you speak a little more to this point in this section?

*straightforward--just a bit of text edit.*

L382: You might consider changing text to “increases with age AT LEAST up to the 100-yr threshold examined here” so that it does not come across as suggesting that NEP only increases for the first hundred years of stand development, which we know is not true.

**Reworded to read, "Notably, net carbon sequestration ($NEP$) exhibits an overall increase with age across the first 100 years of stand development...".**

### 2021-01-24 manuscript/ERL_global_C_review.Rmd
some rearrangmenet in discussion, main comment of R4
- discussion rearrangment suggested by @hmullerlandau (move section on C variable coverage and budget closure to later in the discussion)
- added paragraph on lack of representation of degraded and fragmented forests

*work on this par:*

*ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs. `r flux_age_trends`- `r stock_age_trends`).
The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes.
There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs. `r schematic`, `r `all_diagrams_mature`, `r stock_age_trends`).
Consistent with theory and previous studies (Fig. `r schematic`), the majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests, which are defined by the differences between in- and out- fluxes, do not vary systematically across biomes.
Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Figs. `r age_trends_all`).

*HM: I recommend putting this later in the discussion.  *

There are of course notable holes in the ForC variable coverage (Fig. `r figure_map`) that limit the scope of our inferences here. 
Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Table 1, Figs. `r DW_tot_fig`-`r DW_down_fig`). 
Geographically, all variables are poorly covered in Africa and Siberia, a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
*ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_restructured_2020], or the human footprint in global forests [@magnani_human_2007].

@@

Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
The consistency with which this occurs across numerous fluxes is not surprising, particularly given commonality in the data analyzed or used for calibration, but has never been simultaneously assessed across such a large number of variables [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].

@@

Notably, net carbon sequestration ($NEP$) increases with age up to the 100-yr threshold examined here, with more pronounced patterns in temperate than boreal forests (Fig. `r flux_age_trends`). 
This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig. `r diagram_TeN_mature`).
A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].

@@

@@

## Acknowledgements
Thanks to all researchers whose data are included in *ForC* and this analysis, to Jennifer McGarvey and Ian McGregor for help with the database, and to Norbert Kunert for helpful discussion.
Funding sources included a Smithsonian Scholarly Studies grant to KAT and HML and a Smithsonian Working Land and Seascapes grant to KAT.

>>>>

*ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
Carbon cycling rates generally increased from boreal to tropical regions and with stand age (Figs. `r schematic`, `r all_age_trends`).
Specifically, most C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs. `r schematic`, `r flux_age_trends`- `r stock_age_trends`, `r age_trends_all`).
The notable exception was mature forest $NEP$, which, as the difference between $GPP$ and $R_{eco}$, was indistinguishable across biomes (Fig. `r flux_age_trends`f).
There was also little directional variation in mean mature forest C stocks across biomes, although maximum values for the majority of stocks (all including live or standing woody biomass) occurred in temperate biomes (Figs. `r schematic`, `r `all_diagrams_mature`, `r stock_age_trends`).
Consistent with theory and previous studies (Fig. `r schematic`), the majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests, which are defined by the differences between in- and out- fluxes, do not vary systematically across biomes.
Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C stocks and fluxes (Figs. `r age_trends_all`).

@@

Our analysis reveals that carbon cycling is most rapid in the tropics and slowest in boreal regions, including C fluxes into ($GPP$), within (e.g., $NPP$ and its components), and out of (e.g., $R_{soil}$, $R_{eco}$) the ecosystem.
For mature forests, this is consistent with a large body of previous work demonstrating that C fluxes generally decline with latitude (or increase with temperature) on a global scale [e.g., @luyssaert_co2_2007;  @gillman_latitude_2015; @li_mapping_2019; @banbury_morgan_global_nodate].
This consistency is not not surprising, particularly given commonality in the data analyzed or used for calibration.
The finding that these patterns hold consistently across across numerous fluxes, while unsurprising, is novel to this analysis [but see @banbury_morgan_global_nodate for nine autotrophic fluxes].

@@

Notably, net carbon sequestration ($NEP$) exhibits an overall increase with age across the first 100 years of stand development, with more pronounced patterns in temperate than boreal forests (Fig. `r flux_age_trends`). 
This finding is largely consistent with, but built from a far larger dataset than, previous studies showing an increase in $NEP$ across relatively young stand ages [@pregitzer_carbon_2004; @baldocchi_fluxnet_2001; @luyssaert_old-growth_2008].
However, $NEP$ has been observed to decline from intermediate to old stands [@luyssaert_old-growth_2008], and the $NEP$ estimated by our model for 100-year-old temperate conifer stands (~5 Mg C ha^-1^ yr^-1^) exceeds the mean of mature forests in the same biome (0.7 Mg C ha^-1^ yr^-1^; Fig. `r diagram_TeN_mature`).
A decrease in $NEP$ is consistent with the observed deceleration of biomass accumulation as stands age, although both biomass and non-living C stocks will often continue to increase well beyond the 100-yr threshold used here to delimit young and mature stands [@luyssaert_old-growth_2008; @mcgarvey_carbon_2014; @lichstein_biomass_2009].

@@

There are of course notable holes in the ForC variable coverage that limit the scope of our inferences here. 
Notably, *ForC* currently has sparse--if any--coverage of fluxes to herbivores and higher consumers, along with the woody mortality ($M_{woody}$) and dead wood stocks (Table 1, Figs. `r DW_tot_fig`-`r DW_down_fig`). 
*ForC* does not include soil carbon, which is covered by other efforts [e.g., @kochy_global_2015]. 
*ForC* is not intended to replace databases that are specialized for particular parts of the C cycle analyses, e.g., aboveground biomass [@spawn_harmonized_2020], land-atmosphere fluxes [@baldocchi_fluxnet_2001], soil respiration [@jian_restructured_2020], or the human footprint in global forests [@magnani_human_2007].

Importantly, *ForC* and the analyses presented here cover the forests that have received research attention, which are not a representative sample of the world's existing forests--geographically or in terms of human impacts [@martin_mapping_2012].
Geographically, all variables are poorly covered in Africa and Siberia (Fig. `r figure_map`), a common problem in the carbon-cycle community [@xu_contribution_2016; @schimel_effect_2015]. 
In terms of human impacts, research efforts tend to focus on interior forest ecosystems [@martin_mapping_2012], often in permanently protected areas [e.g., @davies_forestgeo_2021].
Studies of regrowth forests tend to focus on sites where recurring anthropogenic disturbance is not a confounding factor.
Yet, fragmentation and degradation impact a large and growing proportion of Earth's forests [@fao_and_unep_state_2020].
Fragmentation and the creation of edges strongly impacts forest C cycling [e.g., @chaplin-kramer_degradation_2015; @remy_strong_2016; @reinmann_edge_2017; @smith_evidence_2019; @reinmann_urbanization_2020;  @ordway_carbon_2020].
Partial logging and other forms of non- stand clearing anthropogenic disturbance also strongly impact forest C cycling [e.g., @huang_long-term_2010; @piponiot_carbon_2016] but are under-studied [@sist_tropical_2015] and excluded from this analysis. 
Fragmented and degraded forests do not fit the idealized conceptual framework around which this review is structured (Fig. `r schematic`), yet their representation in models, sustainability assessments, and C accounting systems is critical to accurate accounting of C cycling in Earth's forests [e.g., @huang_long-term_2010; @reinmann_edge_2017; @smith_evidence_2019; @piponiot_can_2019]. 
Additional research and synthesis are needed to fill critical gaps in our understanding of forest C cycling [e.g., @sist_tropical_2015].

@@

## Acknowledgements
Thanks to all researchers whose data are included in *ForC* and this analysis, to Jennifer McGarvey and Ian McGregor for help with the database, and to Norbert Kunert for helpful discussion.
Three anonymous reviewers provided helpful feedback. 
Funding sources included a Smithsonian Scholarly Studies grant to KAT and HML and a Smithsonian Working Land and Seascapes grant to KAT.
